---
title: "Computing and making sense of the connectivity maps"
format: html
---

# Computing functionality and flow

This is done via a quarto document with Julia chunks. The outputs of the script
are the functionality and flow maps for each scenario.

# Processing and summarizing the connectivity maps

The next step is to process, visualize, and summarize the habitat and the 
connectivity maps.

## Packages

```{r packages}
library(dplyr)
library(DT)

library(terra)
library(sf)
library(tmap)

library(oneimpact)
library(oneimpactools)
```

## Parameters

```{r parms}
crs_no <- 25833
```

## Connect to PostGIS

```{r connect_pgis}
# This connects to the gisdatabase with a DBI connection named `con`.
# Use for example dbGetQuery(con, "SELECT * FROM ….") to query the database
source("~/.pgpass")

NinaR::postgreSQLConnect(
  host = "gisdata-db.nina.no",
  dbname = "gisdata",
  username = pg_username,
  password = pg_password
)

rm(pg_username, pg_password)
```

## Read scenario table

```{r read_scenario_table}
scen_tab_fil <- "/data/P-Prosjekter/41203800_oneimpact/06_analyses/01_tiltak_MDir_2022/03_output/barriers2/scenarios_mdir_2023_tiltak_rondane_ss1.rds"
scenario_info <- readRDS(file = scen_tab_fil)
```

## Read region - official and practical polygons

```{r get_region}
# get region polygon
region <- sf::st_read(con, 
                      query = db_make_query("SELECT * FROM ",
                                            scenario_info$region_postgis_layer[1],
                                            " WHERE ",
                                            scenario_info$region_postgis_condition[1]))

# read official polygon
region_official <- sf::st_read(con, 
                               query = db_make_query("SELECT * FROM ",
                                                     "sam_wrein_ancillary.reindeer_areas_official_2023",
                                                     " WHERE ",
                                                     "name_area = 'Rondane'"))

# part of the official limits of Rondane within Rondane Nord
region_official <- region_official |> 
  sf::st_intersection(region)

# plot(region_official[1])
```

We also read the barrier, to fill them with value ZERO in the flow maps.

```{r barriers}
# read barriers
barriers <- sf::st_read(con, 
                        query = db_make_query("SELECT * FROM sam_scenarios_wrein.scenarios_mdir_2023_tiltak_rondane_barrier_refined_pol"))
#plot(barriers[1])

# ref rast
ref_rast <- terra::rast(scenario_info$region_file_raster[1]) 

# rasts
barriers_rast <- terra::rasterize(terra::vect(barriers), ref_rast)
barriers_rast <- terra::crop(barriers_rast, region, mask = TRUE)
# plot(barriers_rast)
```

## For one scenario

### Select scenario

```{r select1}
# all scenarios
scenario_info |> 
  dplyr::select(s_gid, contains("scen")) |> 
  DT::datatable()

# select one scenario
selected_scen <- 1
```

### Read maps

```{r read1}
# folder and prefix
dir <- scenario_info$scenario_folder_path[selected_scen]
pre <- scenario_info$scenario_map_prefix[selected_scen]

# files
patt <- c("suit", "perm", "func", "flow")
ff <- list.files(dir, pattern = pre, full.names = TRUE) |> 
  grep(pattern = paste(patt, collapse = "|"), value = TRUE) |> 
  grep(pattern = ".asc$", value = TRUE)
ff <- sapply(patt, function(x) grep(x, ff, value = T))

# read
maps <- terra::rast(ff)
names(maps) <- patt
terra::crs(maps) <- paste0("epsg:", crs_no)
maps <- terra::crop(maps, region, mask = TRUE)

# normalize suit and perm
maps[[1]] <- oneimpact::raster_rescale(maps[[1]])
maps[[2]] <- oneimpact::raster_rescale(maps[[2]])

# sqrt + normalize functionality maps
maps <- c(maps, sqrt(maps[[3]]))
names(maps)[5] <- "func_sqrt"
baseline_max_func_sq <- as.numeric(terra::global(maps[[5]], fun = quantile, prob = 0.975, na.rm = T))
maps[[5]] <- maps[[5]]/baseline_max_func_sq
maps[[5]] <- terra::ifel(maps[[5]] > 1, 1, maps[[5]])
# normalize functionality maps
baseline_max_func <- as.numeric(terra::global(maps[[3]], fun = quantile, prob = 0.975, na.rm = T))
maps[[3]] <- maps[[3]]/baseline_max_func
maps[[3]] <- terra::ifel(maps[[3]] > 1, 1, maps[[3]])

# log + normalize flow maps
maps <- c(maps, log(maps[[4]]))
names(maps)[6] <- "flow_log"
baseline_max_flow_log <- as.numeric(terra::global(maps[[6]], fun = quantile, prob = 0.975, na.rm = T))
maps[[6]] <- oneimpact::raster_rescale(maps[[6]], to = c(0, baseline_max_flow_log))
maps[[6]] <- terra::ifel(maps[[6]] > 1, 1, maps[[6]])
# normalize flow maps
baseline_max_flow_log <- as.numeric(terra::global(maps[[4]], fun = quantile, prob = 0.975, na.rm = T))
maps[[4]] <- maps[[4]]/baseline_max_flow_log
maps[[4]] <- terra::ifel(maps[[4]] > 1, 1, maps[[4]])

# plot(maps)
```

### Plot maps

```{r plot1}
# plot suitability
map1 <- tm_shape(maps[["suit"]]) +
  tm_raster(palette = "Greens", style = "cont", title = "Suitability") +
  tm_layout(legend.position = c("LEFT", "BOTTOM")) +
  tm_shape(region) +
  tm_borders() + 
  tm_compass()
map1

# plot permeability
map2 <- tm_shape(maps[["perm"]]) +
  tm_raster(palette = "-Greys", style = "cont", title = "Permeability") +
  tm_layout(legend.position = c("LEFT", "BOTTOM")) +
  tm_shape(region) +
  tm_borders() + 
  tm_scale_bar(position = c("center", "BOTTOM"))
map2

map3 <- tm_shape(maps[["func"]]) +
  tm_raster(palette = "viridis", style = "cont", title = "Functionality") +
  tm_layout(legend.position = c("LEFT", "BOTTOM"), legend.bg.color = "white") +
  tm_shape(region) +
  tm_borders()
map3

map5 <- tm_shape(maps[["func_sqrt"]]) +
  tm_raster(palette = "viridis", style = "cont", title = "ECA") +
  tm_layout(legend.position = c("LEFT", "BOTTOM"), legend.bg.color = "white") +
  tm_shape(region) +
  tm_borders()
map5

map4 <- tm_shape(maps[["flow"]]) +
  tm_raster(palette = "magma", style = "cont", title = "Flow") +
  tm_layout(legend.position = c("LEFT", "BOTTOM"), legend.bg.color = "white") +
  tm_shape(region) +
  tm_borders()
map4

map6 <- tm_shape(maps[["flow_log"]]) +
  tm_raster(palette = "magma", style = "cont", title = "log Flow") +
  tm_layout(legend.position = c("LEFT", "BOTTOM"), legend.bg.color = "white") +
  tm_shape(region) +
  tm_borders()
map6

plot_list <- list(map1, map2, map3, map5, map4, map6)
names(plot_list) <- c(patt[1:3], "func_sqrt", patt[4], "flow_log") 
tmap::tmap_arrange(map1, map2, map3, map5, map4, map6, ncol = 3, nrow = 2)
```

## For multiple scenarios

### Read maps

Read and normalize maps

```{r read_multiple}
patt <- c("suit", "perm", "func", "func", "func", "flow", "flow")
patt2 <- c("suit", "perm", "func_orig", "func", "func_sqrt", "flow", "flow_log")
scen = nrow(scenario_info)

mask_barriers <- FALSE
dist_decay <- "_50_" # "_500_"

# read
scenario_output <- list()
for(cont in seq_along(patt2)) {
  
  # folder and prefix
  dir <- scenario_info$scenario_folder_path[1]
  pre <- scenario_info$scenario_map_prefix
  
  # files
  pat <- patt[cont]
  pat2 <- patt2[cont]
  ff <- list.files(dir, pattern = pat, full.names = TRUE) |> 
    grep(pattern = paste(pre, collapse = "|"), value = TRUE) |> 
    grep(pattern = ".asc$", value = TRUE)
  if(pat %in% c("func", "flow")) ff <- ff |> grep(pattern = dist_decay, value = T)
  ff <- sort(ff)
  
  # read
  scenario_output[[cont]] <- terra::rast(ff)
  names(scenario_output[[cont]]) <- pre
  names(scenario_output)[cont] <- pat2
  # crs
  terra::crs(scenario_output[[cont]]) <- paste0("epsg:", crs_no)
  # crop and mask
  scenario_output[[cont]] <- terra::crop(scenario_output[[cont]], region, mask = TRUE)
  
  # normalize suitability and permeability maps
  if(pat2 %in% c("suit", "perm")) {
    min_r <- terra::global(scenario_output[[cont]][[1]], fun = "min", na.rm = TRUE)$min
    # max not needed, this was done before exporting
    # max_r <- as.numeric(terra::global(scenario_output[[cont]][[1]], fun = quantile, probs = 0.975, na.rm = TRUE))
    max_r <- terra::global(scenario_output[[cont]][[1]], fun = "max", na.rm = TRUE)$max
    scenario_output[[cont]] <- oneimpact::raster_rescale(scenario_output[[cont]], 
                                                         from = c(min_r, max_r))
  }
  
  # normalize functionality maps
  if(pat2 == "func") {
    baseline_max_func <- terra::global(scenario_output[[cont]][[1]], fun = quantile, prob = 0.975, na.rm = T)[,1]
    scenario_output[[cont]] <- scenario_output[[cont]]/baseline_max_func
    scenario_output[[cont]] <- terra::ifel(scenario_output[[cont]] > 1, 1, scenario_output[[cont]])  
  }
  
  # sqrt + normalize functionality maps
  if(pat2 == "func_sqrt") {
    scenario_output[[cont]] <- sqrt(scenario_output[[cont]])
    baseline_max_func_sq <- terra::global(scenario_output[[cont]][[1]], fun = quantile, prob = 0.975, na.rm = T)[,1]
    scenario_output[[cont]] <- scenario_output[[cont]]/baseline_max_func_sq
    scenario_output[[cont]] <- terra::ifel(scenario_output[[cont]] > 1, 1, scenario_output[[cont]])
  }
  
  # normalize flow maps
  if(pat2 == "flow") {
    if(mask_barriers) scenario_output[[cont]][barriers_rast == 1] <- terra::global(scenario_output[[cont]][[1]], fun = min, na.rm = T)[,1]
    baseline_max_flow <- terra::global(scenario_output[[cont]][[1]], fun = quantile, prob = 0.975, na.rm = T)[,1]
    scenario_output[[cont]] <- scenario_output[[cont]]/baseline_max_flow
    scenario_output[[cont]] <- terra::ifel(scenario_output[[cont]] > 1, 1, scenario_output[[cont]])  
    min_r <- terra::global(scenario_output[[cont]][[1]], fun = "min", na.rm = TRUE)[,1]
    # max_r <- as.numeric(terra::global(scenario_output[[cont]][[1]], fun = "max", na.rm = TRUE))
    scenario_output[[cont]] <- oneimpact::raster_rescale(scenario_output[[cont]], from = c(min_r, 1))
  }
  
  if(pat2 == "flow_log") {
    # log + normalize flow maps
    if(mask_barriers) scenario_output[[cont]][barriers_rast == 1] <- terra::global(scenario_output[[cont]][[1]], fun = min, na.rm = T)[,1]
    scenario_output[[cont]] <- log(scenario_output[[cont]])
    baseline_max_flow_log <- terra::global(scenario_output[[cont]][[1]], fun = quantile, prob = 0.975, na.rm = T)[,1]
    scenario_output[[cont]] <- scenario_output[[cont]]/baseline_max_flow_log
    scenario_output[[cont]] <- terra::ifel(scenario_output[[cont]] > 1, 1, scenario_output[[cont]])
    min_r <- terra::global(scenario_output[[cont]][[1]], fun = "min", na.rm = TRUE)[,1]
    scenario_output[[cont]] <- oneimpact::raster_rescale(scenario_output[[cont]], from = c(min_r, 1))
    
    # scenario_output[[cont]] <- log(scenario_output[[cont]])
    # baseline_max_flow_log <- terra::global(scenario_output[[cont]][[1]], fun = quantile, prob = 0.975, na.rm = T)[,1]
    # scenario_output[[cont]] <- scenario_output[[cont]]/baseline_max_flow_log
    # scenario_output[[cont]] <- terra::ifel(scenario_output[[cont]] > 1, 1, scenario_output[[cont]])
    # min_r <- terra::global(scenario_output[[cont]][[1]], fun = "min", na.rm = TRUE)[,1]
    # scenario_output[[cont]] <- oneimpact::raster_rescale(scenario_output[[cont]], from = c(min_r, 1))
  }
  
  # min_r <- min(terra::global(scenario_output[[cont]], fun = "min", na.rm = TRUE)$min)
  # max_r <- max(terra::global(scenario_output[[cont]], fun = "max", na.rm = TRUE)$max)
  # for(i in 1:nlyr(scenario_output[[cont]])) plot(scenario_output[[cont]][[i]], range = c(min_r, max_r))
  # for(i in 1:nlyr(scenario_output[[cont]])) plot(scenario_output[[cont]][[i]])
}
```

### Plot maps

```{r read_multiple}
patt2 <- c("suit", "perm", "func", "func_sqrt", "flow", "flow_log")
scenario_output_plot <- scenario_output[patt2] # remove original functionality from plots

cont <- 1
plot_list <- list()
# loop over scenarios
for(cont in seq_len(terra::nlyr(scenario_output_plot[[1]]))) {
  
  # plot suitability
  map1 <- tm_shape(scenario_output_plot[[patt2[1]]][[cont]]) +
    tm_raster(palette = "Greens", style = "cont", title = "Suitability") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders() + 
    tm_compass()
  # map1
  
  # plot permeability
  map2 <- tm_shape(scenario_output_plot[[patt2[2]]][[cont]]) +
    tm_raster(palette = "-Greys", style = "cont", title = "Permeability") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders() + 
    tm_scale_bar(position = c("center", "BOTTOM"))
  # map2
  
  # plot flow
  map3 <- tm_shape(scenario_output_plot[[patt2[3]]][[cont]]) +
    tm_raster(palette = "viridis", style = "cont", title = "Functionality") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders()
  # map3
  
  map4 <- tm_shape(scenario_output_plot[[patt2[4]]][[cont]]) +
    tm_raster(palette = "viridis", style = "cont", title = "ECH") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders()
  # map4
  
  map5 <- tm_shape(scenario_output_plot[[patt2[5]]][[cont]]) +
    tm_raster(palette = "magma", style = "cont", title = "Flow") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders()
  # map5
  
  map6 <- tm_shape(scenario_output_plot[[patt2[6]]][[cont]]) +
    tm_raster(palette = "magma", style = "cont", title = "log Flow") +
    tm_layout(legend.position = c("LEFT", "BOTTOM")) +
    tm_shape(region) +
    tm_borders()
  # map6
  
  plot_list <- list(map1, map2, map3, map5, map4, map6)
  names(plot_list) <- scenario_info$scen_name
  print(tmap::tmap_arrange(map1, map2, map3, map5, map4, map6, ncol = 3, nrow = 2))
  # print(tmap::tmap_arrange(map1, map2, ncol = 2, nrow = 1))
}
```

### Save final maps

```{r save_maps}
str(scenario_output_plot, max.level = 2)

# folder
dir <- "/data/P-Prosjekter/41203800_oneimpact/06_analyses/01_tiltak_MDir_2022/03_output/barriers2/final_maps/"
if(!dir.exists(dir)) dir.create(dir)

# save

# loop over layers
for(i in seq_along(scenario_output_plot)) {
  
  # layer
  ly <- names(scenario_output_plot)[i]
  
  # loop over scenarios
  for(j in 1:terra::nlyr(scenario_output_plot[[i]])) {
    
    # name
    name <- scenario_info$scen_nickname[j]
    # map
    kart <- scenario_output_plot[[i]][[j]]
    # save
    f_name <- paste0(dir, "scenario_barriers_", name, "_", ly, dist_decay, ".tif")
    terra::writeRaster(kart, filename = f_name, overwrite = TRUE)
    
  }
}
```


### Summarize changes

```{r summ_change}
patt2 <- c("suit", "perm", "func_orig", "func", "func_sqrt", "flow", "flow_sqrt")
index <- seq_along(patt2)

# crop maps for the official areas
scenario_output_official <- lapply(index, function(i) scenario_output[[i]] |> 
         terra::crop(region_official, mask = TRUE))
names(scenario_output_official) <- names(scenario_output)

# sums
sums <- sapply(index, function(i) terra::global(scenario_output_official[[i]], fun = "sum", na.rm = TRUE))
names(sums) <- names(scenario_output)
sums <- sums[c("suit", "func_orig", "flow")]
sums$func_orig <- sqrt(sums$func_orig) * prod(terra::res(scenario_output_official[[1]])) / 1e6 # km2

#region_official |> sf::st_area() |> (\(x) x/1e6)()

diff_change <- lapply(sums, function(x) x - x[1])
prop_change <- lapply(sums, function(x) (x/x[1] - 1)*100)

summ_tab <- scenario_info |> 
  dplyr::select(s_gid, "project_name", "study_area", "sub_area", contains("scenset_"), contains("scen_")) |> 
  dplyr::mutate(suit_sum = sums[[1]],
                suit_prop_change = prop_change[[1]],
                func_sum = sums[[2]],
                func_prop_change = prop_change[[2]],
                func_km2_prime = diff_change[[2]],
                func_km2_avg = diff_change[[2]]/0.53,
                flow_sum = sums[[3]],
                flow_prop_change = prop_change[[3]])
  # dplyr::mutate(suit = lapply(index, function(x) scenario_output[[x]][[1]])

# export table
summ_tab |> 
  dplyr::select(s_gid, project_name, study_area, sub_area, scen_nickname, 
                suit_prop_change, func_prop_change, flow_prop_change, contains("km2")) |> 
  write.csv(file = paste0("/data/P-Prosjekter/41203800_oneimpact/06_analyses/01_tiltak_MDir_2022/03_output/barriers2/final_maps/summary_table", dist_decay, ".csv"), row.names = FALSE)

summ_tab |> 
  dplyr::select(s_gid, project_name, study_area, sub_area, scen_nickname, 
                suit_prop_change, func_prop_change, flow_prop_change, contains("km2"))
```

